---
title: "Project 3b - Data Analysis and Visualizations"
author: "Sufian"
date: "10/16/2019"
output: html_document
---



Rpub links:

http://rpubs.com/ssufian/540285



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}

# Load packages
library(rvest)
library(stringr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(xml2)
library(httr)
library("tm")
library("SnowballC")
library("wordcloud")
library("RColorBrewer")
library(tidytext)
library("magrittr")
require(ggthemes)


```

## Reading each teammates' job site scrapped job listings

```{r}
#read from Leticia
ltcancel<-read.csv("https://raw.githubusercontent.com/ltcancel/Project3/master/SimplyHiredJobs.csv")
colnames(ltcancel)<-c("Position", "Company","Location","Salary","URL","Job_Description")

#read from Salma
selshahawy<-read.csv("https://raw.githubusercontent.com/salma71/MSDS_2019/master/Fall2019/aquisition_management_607/project_3/jobs_detailsInfo.csv", stringsAsFactors = FALSE) 
colnames(selshahawy)<-c("Position", "Company","Location","URL","Job_Description") 

#read from Sufian
ssufian<-read.csv("https://raw.githubusercontent.com/Luz917/data607project3_ssufian_monster_jobs/master/monsterjobs.csv", stringsAsFactors = FALSE) 
colnames(ssufian)<-c("Position", "Company","Location","Salary","URL","Job_Description")  


```

### Merging all scrapped files into one


```{r }
twocsv<-merge(ltcancel,selshahawy,all= TRUE)

allcsv<-merge(twocsv,ssufian, all=TRUE)

head(allcsv)
```




```{r}
library("magrittr")
# REmove Salary column

clean_allcsv <- allcsv[,-c(4,6)]

head(clean_allcsv)


```


```{r}

# Removed brackets & general cleanup of job summaries

clean_allcsv <- clean_allcsv %>% mutate_each(funs(tolower), Position, Company, Location, Job_Description)

clean_allcsv1 <- str_replace_all(clean_allcsv$Job_Description, "[\r\n]" , "")
clean_allcsv1 <- str_replace_all(clean_allcsv$Job_Description, "\\." , "")
clean_allcsv1 <- str_replace_all(clean_allcsv$Job_Description, "\\*" , "")
clean_allcsv1 <- str_replace_all(clean_allcsv$Job_Description, "\\+" , "")
clean_allcsv1 <- str_replace_all(clean_allcsv$Job_Description, "\\\\" , "")
clean_allcsv1 <- str_replace_all(clean_allcsv$Job_Description, "\\)" , "")
clean_allcsv1 <- str_replace_all(clean_allcsv$Job_Description, "\\(" , "")
clean_allcsv1 <- str_replace_all(clean_allcsv$Job_Description, "[[:punct:]]+", "")
clean_allcsv1 <- str_replace_all(clean_allcsv$Job_Description, "[\n]" , "")


head(clean_allcsv1 )
              
clean_allcsv2 <- clean_allcsv[,-c(4)] # this is for sentinment later


clean_allcsv2$Job_Description <- clean_allcsv1 
  
```



```{r}
#tfidf <- clean_allcsv1
tfidf <- clean_allcsv2$Job_Description 

# Control list to be used for all corpuses
control_list <- list(removePunctuation = TRUE, stopwords = TRUE, tolower = TRUE,
                     weighting = weightTfIdf)

```

```{r}
# TF-IDF Based on the 3 job postings

corpus.all <- VCorpus(VectorSource(tfidf))
tdm.all <- TermDocumentMatrix(corpus.all, control = control_list)
inspect(tdm.all)
# Remove outliers consisting of very rare terms
tdm.60 <- removeSparseTerms(tdm.all, sparse = 0.60)
inspect(tdm.40)
# Sum rows for total & make dataframe
df_all <- tidy(sort(rowSums(as.matrix(tdm.60))))
colnames(df_all) <- c("words", "count")
View(df_all)


```




```{r}


# Graph
ggplot(tail(df_all, 25), aes(reorder(words, count), count)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "TF-IDF of 3 Job Postings",
       x = "Words", y = "Frequency") +
  coord_flip()

```
```{r}
docs <- Corpus(VectorSource(clean_allcsv2$Job_Description))
toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
docs <- tm_map(docs  , toSpace, "/")
docs <- tm_map(docs, toSpace, "@")
docs <- tm_map(docs, toSpace, "\\|")

# Remove numbers
docs <- tm_map(docs, removeNumbers)
# Remove english common stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))
# Remove punctuations
docs <- tm_map(docs, removePunctuation)
# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)
# Text stemming
#docs <- tm_map(docs, stemDocument)

inspect(docs)
dtmw <- TermDocumentMatrix(docs )
m <- as.matrix(dtmw)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq=v)
head(d, 10)

# this is for testing purpose only for now - could be deleted later
#m2 <- as.data.frame(t(m))
#m_wide <- bind_cols(clean_allcsv2, m2)
#m_long <- m_wide %>%
#  gather("word", "count", -Position:-Job_Description)
#View(head(m_long, 1000))


```

### Generate the Word cloud

The important key skill sets can be illustrated as a word cloud as follow :


```{r}
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words=200, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))
```

### Explore frequent terms and their associations


```{r}

findFreqTerms(dtmw, lowfreq = 4)


findAssocs(dtmw, terms = "data scientist", corlimit = 0.3)

```


### Top 10 most frequent words in the job descriptions of the 3 job sites

```{r}
barplot(d[1:10,]$freq, las = 2, names.arg = d[1:10,]$word,
        col ="lightblue", main ="Most frequent words",
        ylab = "Word  frequencies")
```
```{r}



```

### Frequency of Technical skills

```{r}

toolskills <- clean_allcsv2 %>%
    mutate(R = grepl("\\bR\\b,", Job_Description, ignore.case=TRUE)) %>%
    mutate(python = grepl("Python", Job_Description, ignore.case=TRUE)) %>%
    mutate(SQL = grepl("SQL", Job_Description, ignore.case=TRUE)) %>%
    mutate(hadoop = grepl("hadoop", Job_Description, ignore.case=TRUE)) %>%
    mutate(perl = grepl("perl", Job_Description, ignore.case=TRUE)) %>%
    mutate(matplotlib = grepl("matplotlib", Job_Description, ignore.case=TRUE)) %>%
    mutate(Cplusplus = grepl("C++", Job_Description, fixed=TRUE)) %>%
    mutate(VB = grepl("VB", Job_Description, ignore.case=TRUE)) %>%
    mutate(java = grepl("java\\b", Job_Description, ignore.case=TRUE)) %>%
    mutate(scala = grepl("scala", Job_Description, ignore.case=TRUE)) %>%
    mutate(tensorflow = grepl("tensorflow",Job_Description, ignore.case=TRUE)) %>%
    mutate(javascript = grepl("javascript", Job_Description, ignore.case=TRUE)) %>%
    mutate(spark = grepl("spark", Job_Description, ignore.case=TRUE)) %>%
   # mutate(spark = grepl("Hadoop", Job_Description, ignore.case=TRUE)) %>%  
    # mutate(spark = grepl("nosql", Job_Description, ignore.case=TRUE)) %>%  
     # mutate(spark = grepl("spark", Job_Description, ignore.case=TRUE)) %>%  
     # mutate(spark = grepl("sas", Job_Description, ignore.case=TRUE)) %>%  
     # mutate(spark = grepl("excel", Job_Description, ignore.case=TRUE)) %>%  
     # mutate(spark = grepl("aws", Job_Description, ignore.case=TRUE)) %>%  
      #mutate(spark = grepl("azure", Job_Description, ignore.case=TRUE)) %>%  
      mutate(spark = grepl("java", Job_Description, ignore.case=TRUE)) %>%   
      #mutate(spark = grepl("tableau", Job_Description, ignore.case=TRUE)) %>% 
  
  
select(Position, Company, R, python, SQL, hadoop, perl, matplotlib, Cplusplus, VB, java, scala, tensorflow, javascript,spark,java)

```

### Setup tool skills for plotting

```{r}
toolskills2 <- toolskills %>% select(-(1:2)) %>% summarise_all(sum) %>% gather(variable,value) %>% arrange(desc(value))
```

### Visualized the most in-demand tool sKills:

```{r}
ggplot(toolskills2,aes(x=reorder(variable, value), y=value)) + geom_bar(stat='identity',fill="green") + xlab('') + ylab('Frequency') + labs(title='Tool Skills') + coord_flip() + theme_minimal()
```

### Frequency of Hard skills

```{r}


hardskills <- clean_allcsv2 %>%
    mutate(machinelearning = grepl("machine learning", Job_Description, ignore.case=TRUE)) %>%
    mutate(modeling = grepl("model", Job_Description, ignore.case=TRUE)) %>%
    mutate(statistics = grepl("statistics", Job_Description, ignore.case=TRUE)) %>%
    mutate(programming = grepl("programming", Job_Description, ignore.case=TRUE)) %>%
    mutate(quantitative = grepl("quantitative", Job_Description, ignore.case=TRUE)) %>%
    mutate(debugging = grepl("debugging", Job_Description, ignore.case=TRUE)) %>%
    mutate(statistical = grepl("statistical",  Job_Description, ignore.case=TRUE)) %>%
    mutate(regression = grepl("regression",  Job_Description, ignore.case=TRUE)) %>%

  
select(Position, Company, machinelearning, modeling, statistics, programming, quantitative, debugging, statistical, regression)


```

### Setup hard skills for plotting

```{r}

hardskills2 <- hardskills %>% 
               select(-(1:2)) %>% 
               summarise_all(sum) %>% 
               gather(variable,value) %>% 
               arrange(desc(value))
```

### Visualized the most in-demand hard sKills:

```{r}
ggplot(hardskills2,aes(x=reorder(variable, value), y=value)) + 
  geom_bar(stat='identity',fill="yellow") + 
  xlab('') + 
  ylab('Frequency') + 
  labs(title='Hard Skills') + 
  coord_flip() + 
  theme_minimal()

```

### Frequency of Soft skills

```{r}

softskills <- clean_allcsv2 %>%
    mutate(workingremote = grepl("working remote", Job_Description, ignore.case=TRUE)) %>% 
    mutate(communication = grepl("communication", Job_Description, ignore.case=TRUE)) %>%
    mutate(collaborative = grepl("collaborate", Job_Description, ignore.case=TRUE)) %>%
    mutate(creative = grepl("creative", Job_Description, ignore.case=TRUE)) %>%
    mutate(critical = grepl("critical", Job_Description, ignore.case=TRUE)) %>%
    mutate(problemsolving = grepl("problem solving", Job_Description, ignore.case=TRUE)) %>%
    mutate(activelearning = grepl("active learning", Job_Description, ignore.case=TRUE)) %>%
    mutate(hypothesis = grepl("hypothesis", Job_Description, ignore.case=TRUE)) %>%
    mutate(organized = grepl("organize", Job_Description, ignore.case=TRUE)) %>%
    mutate(judgement = grepl("judgement", Job_Description, ignore.case=TRUE)) %>%
    mutate(selfstarter = grepl("self Starter", Job_Description, ignore.case=TRUE)) %>%
    mutate(interpersonalskills = grepl("interpersonal skills", Job_Description, ignore.case=TRUE)) %>%
    mutate(detail = grepl("attention to detail", Job_Description, ignore.case=TRUE)) %>%
    mutate(visualization = grepl("visualization", Job_Description, ignore.case=TRUE)) %>%
    mutate(leadership = grepl("leadership", Job_Description, ignore.case=TRUE)) %>%
    mutate(presentation = grepl("presentation", Job_Description, ignore.case=TRUE)) %>%
      mutate(passion = grepl("passion", Job_Description, ignore.case=TRUE)) %>%
      mutate(research = grepl("research", Job_Description, ignore.case=TRUE)) %>%
      mutate(teamwork = grepl("teamwork", Job_Description, ignore.case=TRUE)) %>%
      mutate(integrity = grepl("integrity", Job_Description, ignore.case=TRUE)) %>%
   mutate(passionate= grepl("passionate", Job_Description, ignore.case=TRUE)) %>%

select(Position, Company, workingremote, communication, collaborative, creative, critical, problemsolving, 
  activelearning, hypothesis, organized, judgement, selfstarter, interpersonalskills, detail, 
  visualization, leadership,presentation,passion,research,teamwork,integrity,passionate)

```



### Setup Soft skills for plotting


```{r}

softskills2 <- softskills %>% 
               select(-(1:2)) %>% 
               summarise_all(sum) %>% 
               gather(variable,value) %>% 
               arrange(desc(value))

```

### Visualized the most in-demand soft sKills:

```{r}

ggplot(softskills2,aes(x=reorder(variable, value), y=value)) + geom_bar(stat='identity',fill="orange") + xlab('') + ylab('Frequency') + labs(title='Soft Skills') + coord_flip() + theme_minimal()

```

### Sentiment Analysis using sentimentr package

- sentiment of sentence within each line of job descriptions


```{r}


install.packages('sentimentr')
library('sentimentr')

scores1 <- sentiment_by(clean_allcsv2$Job_Description, by=NULL) %>%
  pull(ave_sentiment)
head(scores1)

#m_wide2 <- m_wide
clean_allcsv2$sentiment_scores<- scores1
clean_allcsv3 <- clean_allcsv2[,-c(4)] # deleting redundant column
m_wide2 <- clean_allcsv3  %>% 
       select(Company, sentiment_scores)


```


```


```{r}
scores1 <- sentiment_by(clean_allcsv2$Job_Description, by=NULL) %>%
  pull(ave_sentiment)
head(scores1)

#m_wide2 <- m_wide
clean_allcsv2$sentiment_scores<- scores1
clean_allcsv3 <- clean_allcsv2[,-c(4)] # deleting redundant column
head(m_wide2 %>% clean_allcsv2$scores %>% 
       
       select(Company, sentiment_scores))

```

